---
title: An investigation into whether investing in Bitcoin is a wise and prudent decision
subtitle: Competition - bitcoin
author: NikkiSarah
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: false
    number_sections: false
    theme: cosmo
editor_options: 
  chunk_output_type: console
---

## **1. Motivation for the analysis**
The overall purpose of this analysis was to determine whether or not this organisation should invest in Bitcoin. The desired methodology included profiling the historical performance of Bitcoin and comparing it against the stockmarket. The CEO was looking for the answers to the following questions: 

1. How does the performance of Bitcoin compare to the S&P 500 and the price of gold?
2. Based on an analysis of Bitcoin's returns and volatility profile:
   + a) Could it help improve the performance of the organisation's portfolio?
   + b) Could it be used as a hedge against inflation?
3. What could a sample portfolio look like that minimises overall risk and incorporates some or all of these assets? 

**Disclaimer**: I am not a financial professional or overly familiar with how to go about a technical analysis of stock performance. The approach is based on desktop research and so any mistakes in methodology, interpretation and conclusions/recommendations are my own.

## **2. Data and Methods**

### **2.1 Data**
The data provided for the analysis is summarised in Table 2.1 below. Note that it came in three separate files, one for the daily Bitcoin information, one for the S&P500 information and one for the monthly gold and Consumer Price Index (CPI) information.

| Name     | Description                                          | File            |
| -------- | ---------------------------------------------------- | --------------- |
| date     | Daily from 17 September 2014 to 17 November 2021     | bitcoin         |
| date     | Daily from 17 September 2014 to 15 November 2021     | sp500           |
| date     | Monthly from September 2014 to November 2021         | gold_cpi        |
| open     | Price/index level at trading day start               | bitcoin / sp500 |
| high     | Highest price/level reached that day                 | bitcoin / sp500 |
| low      | Lowest price/level reached that day                  | bitcoin / sp500 |
| close    | Price/level at trading day close                     | bitcoin / sp500 |
| volume   | Volume traded that day                               | bitcoin         |
| volume   | Volume of shares making up the index traded that day | sp500           |
| gold_usd | Gold price in USD                                    | gold_cpi        |
| cpi_us   | Inflation (CPI) index for the US                     | gold_cpi        |

The CPI data was publicly-available information obtained from the U.S. Bureau of Labor Statistics (https://www.bls.gov/cpi/).

### **2.2 Methods**

At the outset, the Bitcoin and S&P 500 series were converted to a monthly periodicity. As a result, all the analysis between the different series and benchmarks were comparable. In addition, the analysis primarily focused on the closing prices of Bitcoin and the S&P 500 series.

After the three different data sources were combined into a single dataframe, a preliminary exploratory data analysis (EDA) was performed to determine general trends and patterns in the data. This included the calculation of historical returns on a logged basis.

Next, a more in depth assessment of performance was undertaken, which included a number of different risk to return metrics such as Sharpe's Ratio and the Capital Asset Pricing Model (CAPM).

The final section included using similar metrics to the previous section to determine what combination of Bitcoin and Gold would result in the least risk for the current portfolio. Note that unlike the section analysing individual performance, the portfolio assessment was conducted using the discrete (non-logged) returns. Furthermore, five different portfolios were considered:

- 100% Bitcoin
- 75% Bitcoin and 25% Gold
- 50% Bitcoin and 50% Gold
- 25% Bitcoin and 75% Gold, and
- 100% Gold.

## **3. Results**

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
```

```{r data_import, include=FALSE}
# install libraries not in the default list
install.packages("tidyquant")

# import all the libraries required
#library(GGally) # extension to ggplot2
#library(kableExtra)
library(knitr)
library(tidyquant)
library(tidyverse)

# load the data files
bitcoin <- readr::read_csv('./data/bitcoin-usd.csv')
sp500 <- readr::read_csv('./data/sp500.csv')
gold_cpi <- readr::read_csv('./data/monthly_data.csv')

# and take a look at them
summary(bitcoin)
summary(sp500)
summary(gold_cpi)
```

``` {r monthly_conversion_and_aggregation}
# add identifier columns
bitcoin$name <- "bitcoin"
sp500$name <- "sp500"

# combine bitcoin and sp500 into one dataframe
bitcoin_sp500 <- rbind(bitcoin, sp500)

# divide volumes by 1 million
bitcoin_sp500 <- bitcoin_sp500 %>% 
  mutate(volume_m = volume / 1000000)

# convert into a monthly series
bitcoin_sp500_monthly <- bitcoin_sp500 %>% 
  group_by(name) %>% 
  tq_transmute(mutate_fun = to.monthly, indexAt = 'firstof')

# convert gold_cpi into tidy format
# rename the values column to "price" as will be appending to bitcoin_sp500 dataframe
gold_cpi_tidy <- pivot_longer(gold_cpi, cols = gold_usd:cpi_us, 
                              names_to = "name", values_to = "close")

# create extra dummy columns
gold_cpi_tidy <- gold_cpi_tidy %>% 
  mutate(open = NA,
         high = NA,
         low = NA,
         volume = NA,
         volume_m = NA)

# append the tidy dataframe to bitcoin_sp500
asset_df <- rbind(bitcoin_sp500_monthly, gold_cpi_tidy) %>% 
  arrange(name, date)
```

``` {r prices_and_volumes}
# reindex the closing price column to 100
asset_df <- asset_df %>% 
  group_by(name) %>% 
  mutate(init_close = if_else(
    date == min(date), 
    close, 
    NA_real_)) %>% 
  mutate(rebased_close = round(100 * close / sum(init_close, na.rm = TRUE), 1)) %>% 
  ungroup()

# add some performance information
asset_df <- asset_df %>% 
  group_by(name) %>% 
  tq_mutate(select = rebased_close,
            mutate_fun = volatility,
            col_rename = "volatility")

asset_df <- asset_df %>% 
  mutate(facet = if_else(
    name == "bitcoin", 
    1, 
    if_else(
      name == "cpi_us",
      3,
      2
    )))
```

### **3.1 Exploratory Data Analysis**

Figure 3.1 below indicates that the closing price of Bitcoin rose significantly between September 2014 and November 2021, and had a far higher endpoint than any of the other assets under investigation.

``` {r closing_prices_plot}
# plot the rebased closing prices in logged form
p1 <- asset_df %>% 
  ggplot(aes(x = date, y = rebased_close, group = name, colour = name)) +
  scale_y_log10() +
  geom_line() +
  scale_colour_brewer(palette = "Paired",
                      labels = c("Bitcoin", "CPI (US)", "Gold", "S&P 500")) +
  labs(title = "Figure 3.1: Monthly closing prices",
       subtitle = "1 September 2014 to 1 November 2021",
       caption = "Note: Prices were rebased to 1 September 2014 for easier comparison and plotted on a log scale.",
       x = "",
       y = "Logged month-end closing price (1-Sep-14 = 100)",
       colour = "Asset") +
  facet_wrap(vars(facet), ncol = 3, scales = "free") +
  theme_classic()
p1
```

However, Figure 3.2 below indicates that there was also a reasonable amount of closing price volatility in Bitcoin over the same time period. In contrast, gold did not appear to exhibit the same amount of volatility and the CPI was the most stable.

```{r volatilty_plot}
# plot close-to-close volatility
p2 <- asset_df %>% 
  ggplot(aes(x = date, y = volatility, group = name, colour = name)) +
  geom_line() +
  scale_colour_brewer(palette = "Paired",
                      labels = c("Bitcoin", "CPI (US)", "Gold", "S&P 500")) +
  labs(title = "Figure 3.2: Monthly close-to-close price volatility",
       subtitle = "1 September 2014 to 1 November 2021",
       caption = "Note: Volatility was calculated on the rebased prices not the original prices.",
       x = "",
       y = "Month-end closing price",
       colour = "Asset") +
  facet_wrap(vars(facet), ncol = 3, scales = "free") +
  theme_classic()
p2
```

A plot of the monthly volumes in Bitcoin indicates that there has been an almost exponential increase in the amount traded, whereas the monthly volume data for the S&P 500 was relatively stable (Figure 3.3). Unfortunately volume data was not available for gold, which would have been a useful comparator to the Bitcoin series.

```{r volumes_plot}
# plot the monthly volumes
p3 <- asset_df %>% 
  filter(name == "bitcoin" | name == "sp500") %>% 
  ggplot(aes(x = date, y = volume_m)) +
  geom_segment(aes(xend = date, yend = 0, colour = volume_m)) +
  geom_smooth(method = "loess", se = FALSE, size = 0.5) +
  labs(title = "Figure 3.3: Monthly bitcoin and S&P 500 volumes",
       subtitle = "1 September 2014 to 1 November 2021",
       x = "",
       y = "Volume (in millions)") +
  facet_wrap(vars(name), ncol = 2, scales = "free") +
  theme_classic() +
  theme(legend.position = "none")
p3
```

Calculation and plotting of the distribution of the monthly logged returns were largely consistent with the previous charts. That is, that the returns of Bitcoin exhibited the largest variation and the returns from gold were similar to what would be obtained if one could invest directly in the S&P 500. (See Figure 3.4). All assets had logged returns approximately centred around zero.

```{r log_returns}
# now add monthly returns to the dataframe
# logged returns are calculated for structural reasons
asset_df <- asset_df %>% 
  tq_mutate(select = close,
            mutate_fun = periodReturn,
            type = "log",
            col_rename = "logged_return")

# replace any zero returns with NA's (these are just at the start of each series)
asset_df['logged_return'][asset_df['logged_return'] == 0] <- NA

# plot the distributions of the returns
# histograms not plotted
p4 <- asset_df %>% 
  ggplot(aes(x = logged_return)) +
  geom_histogram(colour = "black", fill = "#00CC99") +
  labs(title = "Figure x.x: Distribution of monthly logged returns by asset",
     subtitle = "1 September 2014 to 1 November 2021",
     x = "Logged return",
     y = "Count") +
  facet_wrap(vars(name), ncol = 2, scales = "free") +
  theme_classic()

p5 <- asset_df %>% 
  ggplot(aes(x = name, y = logged_return, fill = name)) +
  geom_violin(scale = "width") +
  geom_boxplot(width = 0.1) +
  geom_jitter(aes(colour = name), shape = 16, position = position_jitter(0.1)) +
  guides(fill = "none") +
  guides(colour = "none") +
  labs(title = "Figure 3.4: Distribution of monthly logged returns by asset",
  subtitle = "1 September 2014 to 1 November 2021",
  x = "Asset",
  y = "Logged return") +
  scale_color_brewer(palette = "Greens") +
  scale_fill_brewer(palette = "Paired") +
  theme_classic()
p5
```

### **3.2 Individual performance assessment**

Indeed, calculation of some basic descriptive statistics on the logged returns indicated that Bitcoin's returns were far higher than either the S&P 500 or Gold (0.06, 0.01 and 0.004 respectively) but Bitcoin carried a commensurate amount of risk (0.2). In contrast, Gold underperformed on average compared to the S&P 500, but had a similar amount of volatility (both 0.04 to 2 decimal places). See Table 3.1 and Figure 3.5 below.

``` {r summary_stats}
# calculate summary statistics for the monthly log returns
log_return_df <- asset_df %>% 
  select(c(name, date, logged_return))

summary_stats <- log_return_df %>% 
  tq_performance(Ra = logged_return, performance_fun = table.Stats) %>% 
  select(name, ArithmeticMean, Maximum, Median, Minimum, Quartile1, Quartile3, 
         Stdev, Variance)
kable(summary_stats, caption = "Table 3.1: Asset summary performance statistics")

# plot of the risk (sd) vs return (mean) values
p6 <- summary_stats %>% 
  ggplot(aes(x = Stdev, y = ArithmeticMean)) +
  geom_point(aes(colour = name), size = 5) +
  labs(title = "Figure 3.5: Logged monthly returns risk vs return",
  subtitle = "1 September 2014 to 1 November 2021",
  x = "Risk (Standard deviation)",
  y = "Return (Mean)",
  colour = "Asset") +
  scale_colour_brewer(palette = "Paired",
                      labels = c("Bitcoin", "CPI (US)", "Gold", "S&P 500")) +
  theme_classic()
p6
```

Calculation of the annualised returns and volatility only confirmed the previous insights (Table 3.2 below). Between September 2014 and November 2021, Bitcoin had an annualised log return of 0.5, compared to 0.04 for Gold, 0.1 for the S&P 500 and 0.02 for the CPI. Similarly, the annualised standard deviation over the same period for Bitcoin was 0.7, 0.1 for Gold and the S&P 500 and 0.01 for the CPI.

The Sharpe ratio is defined as the return per unit of risk (represented by variability). In the classic case (which is what is used in this case), the unit of risk is the standard deviation of the returns. The figures shown in Table 3.3 below indicate that on a risk-adjusted basis, Bitcoin is still a reasonable investment, but Gold is not. However, the Sharpe Ratio indicates that neither Bitcoin nor Gold would be useful as an inflation hedge (presumably because Bitcoin carries too much risk relative to its returns and Gold's returns are too low).

```{r annualised_returns_and_Sharpe}
# annualised returns
annualised_returns <- log_return_df %>%   
    tq_performance(Ra = logged_return, performance_fun = table.AnnualizedReturns)
kable(annualised_returns, caption = "Table 3.2: Asset annualised return figures")

# Sharpe Ratio
Sharpe_Ratio <- log_return_df %>% 
  tq_performance(Ra = logged_return, performance_fun = SharpeRatio, 
                 FUN = "StdDev")
kable(Sharpe_Ratio, caption = "Table 3.3: Asset Sharpe Ratios")
```

The Capital Asset Pricing Model (CAPM) is an excess-return single factor model that relates the return of an asset to a benchmark or baseline return. As such, it describes the relationship between the expected return of an asset and its exposure to systematic risk. The positive Alpha values indicate that both Bitcoin and Gold outperformed the S&P 500, although again Bitcoin was the better performer. Also consistent with previous insights were the Beta values, which indicated that Bitcoin was a riskier investment than the S&P 500, but Gold carried lower market risk and was therefore safer. 

The CAPM model with the CPI as the baseline series also indicated that the use of Bitcoin as an inflation hedge should be considered with caution (positive Alpha indicating overperformance but high Beta indicating high market risk).

Finally, the Information Ratio is a variation on the Sharpe Ratio, adjusting for the market return. It should perhaps be unsurprising by now to note that the Information Ratio indicated that Bitcoin overperformed the S&P 500 and Gold underperformed (0.6 and -0.4 respectively). Both assets performed better than the CPI, but again Bitcoin moreso.

``` {r CAPM}
# create the benchmark dataframes
sp500_benchmark_df <- log_return_df %>% 
  pivot_wider(names_from = "name",
              values_from = "logged_return") %>% 
  pivot_longer(cols = bitcoin:gold_usd, values_to = "Ra") %>% 
  rename(Rb = sp500) %>% 
  arrange(name, date) %>% 
  select(date, name, Ra, Rb)

cpi_benchmark_df <- log_return_df %>% 
  pivot_wider(names_from = "name",
              values_from = "logged_return") %>% 
  pivot_longer(cols = c(bitcoin, gold_usd, sp500), values_to = "Ra") %>% 
  rename(Rb = cpi_us) %>% 
  arrange(name, date) %>% 
  select(date, name, Ra, Rb)

# CAPM
CAPM_sp500 <- sp500_benchmark_df %>% 
  group_by(name) %>% 
  tq_performance(Ra = Ra, Rb = Rb, performance_fun = table.CAPM) %>% 
  select(name, Alpha, Beta, InformationRatio) %>% 
  filter(name != "cpi_us")
kable(CAPM_sp500, caption = "Table 3.4: CAPM statistics using the S&P 500 as the benchmark")

CAPM_cpi <- cpi_benchmark_df %>% 
  group_by(name) %>% 
  tq_performance(Ra = Ra, Rb = Rb, performance_fun = table.CAPM) %>% 
  select(name, Alpha, Beta, InformationRatio) %>% 
  filter(name != "sp500")
kable(CAPM_cpi, caption = "Table 3.5: CAPM statistics using the CPI as the benchmark")
```

### **3.3 Portfolio Investigation**

Since the primary purpose of holding a portfolio of assets is risk diversification, it was prudent to first check the relationship between the returns of the two assets under consideration (Gold and Bitcoin - direct investment in the CPI or the S&P 500 is impossible). Luckily the correlation was reasonably weak and in fact negative at -0.19. (Note that the correlations changed only marginally if the discrete or non-logged returns were used.)

Tables 3.5 and 3.6 display the results of the performance and risk analysis of the five different portfolios considered. Unsurprisingly, the CAPM model indicated that a portfolio consisting of 100% Bitcoin was the best performer relative to the S&P 500, but also carried the highest risk. Choosing to invest solely in Gold was likely to produce only marginally better returns than the S&P 500, but carried near-zero risk. The Information and Sharpe Ratio results were consisted with these observations.

``` {r return_correlations, results='hide'}
# plot the correlation between all the assets
corr_data <- asset_df %>% 
  tq_mutate(select = close,
            mutate_fun = periodReturn,
            col_rename = "return") %>% 
  select(c(name, date, return, logged_return)) %>% 
  pivot_wider(id_cols = c(name, date), names_from = name, 
              values_from = c(return, logged_return)) %>% 
  drop_na()

cor(corr_data$logged_return_bitcoin, corr_data$logged_return_gold_usd)
cor(corr_data$return_bitcoin, corr_data$return_gold_usd)

#corr_plot_returns <- chart.Correlation(corr_data[,2:5])
#corr_plot_log_returns <- chart.Correlation(corr_data[,6:9])
```

```{r portfolio_creation}
# create the asset and benchmark returns dataframes
returns_df <- asset_df %>% 
  tq_mutate(select = close,
            mutate_fun = periodReturn,
            col_rename = "Ra") %>% 
  select(c(name, date, Ra)) %>% 
  filter(name != "cpi_us") %>% 
  filter(name != "sp500")

benchmark_df <- asset_df %>% 
  tq_mutate(select = close,
            mutate_fun = periodReturn,
            col_rename = "Rb") %>% 
  select(c(name, date, Rb)) %>% 
  filter(name == "sp500")

# 'grow' the asset dataframe by the number of portfolios desired
returns_df_multi <- returns_df %>%
    tq_repeat_df(n = 5)

# create the table of weights
weights <- c(
    1, 0,
    0.75, 0.25,
    0.5, 0.5,
    0.25, 0.75,
    0, 1
)
assets <- c("bitcoin", "gold_usd")
weights_table <-  tibble(assets) %>%
    tq_repeat_df(n = 5) %>%
    bind_cols(tibble(weights)) %>%
    group_by(portfolio)

# aggregate the portfolio
portfolio_returns_df_multi <- returns_df_multi %>%
    tq_portfolio(assets_col  = name, 
                 returns_col = Ra, 
                 weights     = weights_table, 
                 col_rename  = "Ra")

# join with the benchmark data
portfolio_df <- left_join(portfolio_returns_df_multi, 
                          benchmark_df,
                          by = "date") %>% 
  select(-name)
```

```{r portfolio_performance_assessment}
portfolio_CAPM <- portfolio_df %>% 
  group_by(portfolio) %>% 
  tq_performance(Ra = Ra, Rb = Rb, performance_fun = table.CAPM) %>% 
  select(portfolio, Alpha, Beta, InformationRatio)
kable(portfolio_CAPM, caption = "Table 3.6: Portfolio CAPM statistics using the S&P 500 as the benchmark")

portfolio_Sharpe <- portfolio_df %>% 
  group_by(portfolio) %>% 
  tq_performance(Ra = Ra, Rb = NULL, performance_fun = SharpeRatio, 
                 FUN = "StdDev")
kable(portfolio_Sharpe, caption = "Table 3.7: Portfolio Sharpe Ratios")
```

### Insights and Recommendations

1. Bitcoin is a high return-high risk investment but historically has performed better than the S&P 500.
2. Gold is at the opposite end of the spectrum, being a low risk-low return investment.
3. Based on Bitcoin's historical performance relative to the S&P 500 and the CPI, it could improve the organisation's current investment portfolio and could be used as an inflation hedge.
4. A risk-minimising portfolio would only include Gold. However, in order to get better returns, some proportion of Bitcoin should also be considered.
5. The analysis should be re-run using adjusted asset prices, which account for any asset splits that may have occurred. Asset splits tend to cause a large decline in stock performance, which penalises the stocks that split by increasing the standard deviation of the daily returns.

### Appendix

#### Appendix A - Code

```{r all_code, eval=FALSE, echo=TRUE}
# install libraries not in the default list
install.packages("tidyquant")

# import all the libraries required
library(knitr)
library(tidyquant)
library(tidyverse)

# load the data files
bitcoin <- readr::read_csv('./data/bitcoin-usd.csv')
sp500 <- readr::read_csv('./data/sp500.csv')
gold_cpi <- readr::read_csv('./data/monthly_data.csv')

# and take a look at them
summary(bitcoin)
summary(sp500)
summary(gold_cpi)

# add identifier columns
bitcoin$name <- "bitcoin"
sp500$name <- "sp500"

# combine bitcoin and sp500 into one dataframe
bitcoin_sp500 <- rbind(bitcoin, sp500)

# divide volumes by 1 million
bitcoin_sp500 <- bitcoin_sp500 %>% 
  mutate(volume_m = volume / 1000000)

# convert into a monthly series
bitcoin_sp500_monthly <- bitcoin_sp500 %>% 
  group_by(name) %>% 
  tq_transmute(mutate_fun = to.monthly, indexAt = 'firstof')

# convert gold_cpi into tidy format
# rename the values column to "price" as will be appending to bitcoin_sp500 dataframe
gold_cpi_tidy <- pivot_longer(gold_cpi, cols = gold_usd:cpi_us, 
                              names_to = "name", values_to = "close")

# create extra dummy columns
gold_cpi_tidy <- gold_cpi_tidy %>% 
  mutate(open = NA,
         high = NA,
         low = NA,
         volume = NA,
         volume_m = NA)

# append the tidy dataframe to bitcoin_sp500
asset_df <- rbind(bitcoin_sp500_monthly, gold_cpi_tidy) %>% 
  arrange(name, date)

# reindex the closing price column to 100
asset_df <- asset_df %>% 
  group_by(name) %>% 
  mutate(init_close = if_else(
    date == min(date), 
    close, 
    NA_real_)) %>% 
  mutate(rebased_close = round(100 * close / sum(init_close, na.rm = TRUE), 1)) %>% 
  ungroup()

# add some performance information
asset_df <- asset_df %>% 
  group_by(name) %>% 
  tq_mutate(select = rebased_close,
            mutate_fun = volatility,
            col_rename = "volatility")

asset_df <- asset_df %>% 
  mutate(facet = if_else(
    name == "bitcoin", 
    1, 
    if_else(
      name == "cpi_us",
      3,
      2
    )))

# plot the rebased closing prices in logged form
p1 <- asset_df %>% 
  ggplot(aes(x = date, y = rebased_close, group = name, colour = name)) +
  scale_y_log10() +
  geom_line() +
  scale_colour_brewer(palette = "Paired",
                      labels = c("Bitcoin", "CPI (US)", "Gold", "S&P 500")) +
  labs(title = "Figure 3.1: Monthly closing prices",
       subtitle = "1 September 2014 to 1 November 2021",
       caption = "Note: Prices were rebased to 1 September 2014 for easier comparison and plotted on a log scale.",
       x = "",
       y = "Logged month-end closing price (1-Sep-14 = 100)",
       colour = "Asset") +
  facet_wrap(vars(facet), ncol = 3, scales = "free") +
  theme_classic()
p1

# plot close-to-close volatility
p2 <- asset_df %>% 
  ggplot(aes(x = date, y = volatility, group = name, colour = name)) +
  geom_line() +
  scale_colour_brewer(palette = "Paired",
                      labels = c("Bitcoin", "CPI (US)", "Gold", "S&P 500")) +
  labs(title = "Figure 3.2: Monthly close-to-close price volatility",
       subtitle = "1 September 2014 to 1 November 2021",
       caption = "Note: Volatility was calculated on the rebased prices not the original prices.",
       x = "",
       y = "Month-end closing price",
       colour = "Asset") +
  facet_wrap(vars(facet), ncol = 3, scales = "free") +
  theme_classic()
p2

# plot the monthly volumes
p3 <- asset_df %>% 
  filter(name == "bitcoin" | name == "sp500") %>% 
  ggplot(aes(x = date, y = volume_m)) +
  geom_segment(aes(xend = date, yend = 0, colour = volume_m)) +
  geom_smooth(method = "loess", se = FALSE, size = 0.5) +
  labs(title = "Figure 3.3: Monthly bitcoin and S&P 500 volumes",
       subtitle = "1 September 2014 to 1 November 2021",
       x = "",
       y = "Volume (in millions)") +
  facet_wrap(vars(name), ncol = 2, scales = "free") +
  theme_classic() +
  theme(legend.position = "none")
p3

# now add monthly returns to the dataframe
# logged returns are calculated for structural reasons
asset_df <- asset_df %>% 
  tq_mutate(select = close,
            mutate_fun = periodReturn,
            type = "log",
            col_rename = "logged_return")

# replace any zero returns with NA's (these are just at the start of each series)
asset_df['logged_return'][asset_df['logged_return'] == 0] <- NA

# plot the distributions of the returns
# histograms not plotted
p4 <- asset_df %>% 
  ggplot(aes(x = logged_return)) +
  geom_histogram(colour = "black", fill = "#00CC99") +
  labs(title = "Figure x.x: Distribution of monthly logged returns by asset",
     subtitle = "1 September 2014 to 1 November 2021",
     x = "Logged return",
     y = "Count") +
  facet_wrap(vars(name), ncol = 2, scales = "free") +
  theme_classic()

p5 <- asset_df %>% 
  ggplot(aes(x = name, y = logged_return, fill = name)) +
  geom_violin(scale = "width") +
  geom_boxplot(width = 0.1) +
  geom_jitter(aes(colour = name), shape = 16, position = position_jitter(0.1)) +
  guides(fill = "none") +
  guides(colour = "none") +
  labs(title = "Figure 3.4: Distribution of monthly logged returns by asset",
  subtitle = "1 September 2014 to 1 November 2021",
  x = "Asset",
  y = "Logged return") +
  scale_color_brewer(palette = "Greens") +
  scale_fill_brewer(palette = "Paired") +
  theme_classic()
p5

# calculate summary statistics for the monthly log returns
log_return_df <- asset_df %>% 
  select(c(name, date, logged_return))

summary_stats <- log_return_df %>% 
  tq_performance(Ra = logged_return, performance_fun = table.Stats) %>% 
  select(name, ArithmeticMean, Maximum, Median, Minimum, Quartile1, Quartile3, 
         Stdev, Variance)
kable(summary_stats, caption = "Table 3.1: Asset summary performance statistics")

# plot of the risk (sd) vs return (mean) values
p6 <- summary_stats %>% 
  ggplot(aes(x = Stdev, y = ArithmeticMean)) +
  geom_point(aes(colour = name), size = 5) +
  labs(title = "Figure 3.5: Logged monthly returns risk vs return",
  subtitle = "1 September 2014 to 1 November 2021",
  x = "Risk (Standard deviation)",
  y = "Return (Mean)",
  colour = "Asset") +
  scale_colour_brewer(palette = "Paired",
                      labels = c("Bitcoin", "CPI (US)", "Gold", "S&P 500")) +
  theme_classic()
p6

# annualised returns
annualised_returns <- log_return_df %>%   
    tq_performance(Ra = logged_return, performance_fun = table.AnnualizedReturns)
kable(annualised_returns, caption = "Table 3.2: Asset annualised return figures")

# Sharpe Ratio
Sharpe_Ratio <- log_return_df %>% 
  tq_performance(Ra = logged_return, performance_fun = SharpeRatio, 
                 FUN = "StdDev")
kable(Sharpe_Ratio, caption = "Table 3.3: Asset Sharpe Ratios")

# create the benchmark dataframes
sp500_benchmark_df <- log_return_df %>% 
  pivot_wider(names_from = "name",
              values_from = "logged_return") %>% 
  pivot_longer(cols = bitcoin:gold_usd, values_to = "Ra") %>% 
  rename(Rb = sp500) %>% 
  arrange(name, date) %>% 
  select(date, name, Ra, Rb)

cpi_benchmark_df <- log_return_df %>% 
  pivot_wider(names_from = "name",
              values_from = "logged_return") %>% 
  pivot_longer(cols = c(bitcoin, gold_usd, sp500), values_to = "Ra") %>% 
  rename(Rb = cpi_us) %>% 
  arrange(name, date) %>% 
  select(date, name, Ra, Rb)

# CAPM
CAPM_sp500 <- sp500_benchmark_df %>% 
  group_by(name) %>% 
  tq_performance(Ra = Ra, Rb = Rb, performance_fun = table.CAPM) %>% 
  select(name, Alpha, Beta, InformationRatio) %>% 
  filter(name != "cpi_us")
kable(CAPM_sp500, caption = "Table 3.4: CAPM statistics using the S&P 500 as the benchmark")

CAPM_cpi <- cpi_benchmark_df %>% 
  group_by(name) %>% 
  tq_performance(Ra = Ra, Rb = Rb, performance_fun = table.CAPM) %>% 
  select(name, Alpha, Beta, InformationRatio) %>% 
  filter(name != "sp500")
kable(CAPM_cpi, caption = "Table 3.5: CAPM statistics using the CPI as the benchmark")

# plot the correlation between all the assets
corr_data <- asset_df %>% 
  tq_mutate(select = close,
            mutate_fun = periodReturn,
            col_rename = "return") %>% 
  select(c(name, date, return, logged_return)) %>% 
  pivot_wider(id_cols = c(name, date), names_from = name, 
              values_from = c(return, logged_return)) %>% 
  drop_na()

cor(corr_data$logged_return_bitcoin, corr_data$logged_return_gold_usd)
cor(corr_data$return_bitcoin, corr_data$return_gold_usd)

corr_plot_returns <- chart.Correlation(corr_data[,2:5])
corr_plot_log_returns <- chart.Correlation(corr_data[,6:9])

# create the asset and benchmark returns dataframes
returns_df <- asset_df %>% 
  tq_mutate(select = close,
            mutate_fun = periodReturn,
            col_rename = "Ra") %>% 
  select(c(name, date, Ra)) %>% 
  filter(name != "cpi_us") %>% 
  filter(name != "sp500")

benchmark_df <- asset_df %>% 
  tq_mutate(select = close,
            mutate_fun = periodReturn,
            col_rename = "Rb") %>% 
  select(c(name, date, Rb)) %>% 
  filter(name == "sp500")

# 'grow' the asset dataframe by the number of portfolios desired
returns_df_multi <- returns_df %>%
    tq_repeat_df(n = 5)

# create the table of weights
weights <- c(
    1, 0,
    0.75, 0.25,
    0.5, 0.5,
    0.25, 0.75,
    0, 1
)
assets <- c("bitcoin", "gold_usd")
weights_table <-  tibble(assets) %>%
    tq_repeat_df(n = 5) %>%
    bind_cols(tibble(weights)) %>%
    group_by(portfolio)

# aggregate the portfolio
portfolio_returns_df_multi <- returns_df_multi %>%
    tq_portfolio(assets_col  = name, 
                 returns_col = Ra, 
                 weights     = weights_table, 
                 col_rename  = "Ra")

# join with the benchmark data
portfolio_df <- left_join(portfolio_returns_df_multi, 
                          benchmark_df,
                          by = "date") %>% 
  select(-name)

portfolio_CAPM <- portfolio_df %>% 
  group_by(portfolio) %>% 
  tq_performance(Ra = Ra, Rb = Rb, performance_fun = table.CAPM) %>% 
  select(portfolio, Alpha, Beta, InformationRatio)
kable(portfolio_CAPM, caption = "Table 3.6: Portfolio CAPM statistics using the S&P 500 as the benchmark")

portfolio_Sharpe <- portfolio_df %>% 
  group_by(portfolio) %>% 
  tq_performance(Ra = Ra, Rb = NULL, performance_fun = SharpeRatio, 
                 FUN = "StdDev")
kable(portfolio_Sharpe, caption = "Table 3.7: Portfolio Sharpe Ratios")
```